---
title: "Reproducible document"
author: "MarijaP"
format: html
---

## Basics of R

```{r}
1 + 1
```

## Setup code chunk


## About me

-Marija Petkovic
-Steno Diabetes Center Copenhagen
-Translational Research

DDEA postdoc working on a **diabetic** *foot complications*

## Simple code
```{r}
3 * 3
```

## Setup code chunk

```{r setup}
# the two options of what to show in the output file (elated to chapter 8 of the course)
# press ctrl-shift-render after you give eg.commands bellow
# this has to be right after r setup code chunk with no space
#| message: false
#|warning: false
library(tidyverse)
post_meal_data <- read_csv2(
  here::here("data/post-meal-insulin.csv")
)
```
## Showing the data
```{r}
post_meal_data
```

```{r}
glimpse(post_meal_data)
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

@fig-bmi-histo

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
#| fig-cap: ""
#| label: fig-istribution_BMI
ggplot(post_meal_data, aes(x = ))
```

See @fig-___ above for a cool plot!

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Putting two plots side by side

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

# the next chunk combines the two graphs

```{r side-by-side-figs}
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#| - "Distribution of BMI"
#| - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting two continuous variables

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point()
```

```{r}
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

```{r}
#| fig-cap: "scatterplot showing BMI agains AUC of blood glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

usethis::gh_token_help()
usethis::create_github_token()

# personal token name
#LearningR

